---
title: "Mid-term Exam #2"
subtitle: "Due date: 6 November"
format: 
  html:
    self-contained: true
toc: true
editor: visual
execute: 
  echo: true
  message: false
  warning: false
---

Please read all of the questions carefully and follow all instructions. Each question has an allotted point value. Be as thorough and specific as possible; extra calculations and incorrect information, even in the presence of correct information, will result in point deductions. Be sure to show all formulas and all necessary work to answer the questions. You may upload your completed exam to the Elms course site (attach to Midterm Exam #2).

::: callout-note
While this is an open-note exam, you are not to receive assistance from anyone else (as usual, the Honor Code applies).
:::

**Total points: 50 points**

## Conceptual questions

::: callout-note
Please include all work (and computations) necessary to answer the questions.
:::

**Total points: 26**

```{r}
library(tidyverse) 
library(skimr) 
library(janitor) 
library(ggsci) 
library(MetBrewer)
library(devtools)
library(polisciols)
library(modelsummary)
library(readr)
library(broom)
library(infer)
library(scales)

```

### Question 1

*2 points*

::: callout-note
1/2 You need to demonstrate how you calculate the critical t-statistic.
:::

Suppose I am interested in determining if freshman undergraduates at the University of Maryland spend more hours studying in the average week than sophomore undergraduates. I conduct a study in which I take a simple random sample (SRS) of 1200 freshman students and 1200 sophomore students. I find that the freshman students in my sample study for, on average, 412 minutes per week and the sophomore students in my sample study for, on average, 335 minutes per week. The standard error of the difference is 30. What is a 90% confidence interval for the difference between freshman and sophomore students?

```{r}
(412 - 335) + 1.645*30

(412 - 335) - 1.645*30

```

The answer for the 90% confidence interval is \[27.65, 126.35\] \### Question 2

*2 points*

::: callout-note
1/2 **You used `pnorm` (Z-distribution)** to calculate the $p$-value for a two-sample mean hypothesis test. Since the population standard deviation ($\sigma$) is unknown (you only have the standard error of the difference, which is computed using sample standard deviations), you must use the **t-distribution** for the $p$-value and critical value comparison.
:::

Based on the results of my study described in question 1, can you reject the null hypothesis of no statistically meaningful difference in the study habits of sophomore and freshman students? Why or why not?

```{r}
2 * pnorm((412-335)/30, lower.tail = F)
```

From the two tail t-test with p-value of 0.01, we can reject the null hypothesis of no statistically meaningful difference. \### Question 3

*2 points*

::: callout-note
2/2
:::

If I am testing a null hypothesis that X has no effect on Y in the population (and thus my alternative hypothesis is that X does have an effect), why might I prefer to commit a Type-II error over a Type-I error (and, of course, this holds aside my first preference of making no error at all)? Answer in no more than two sentences.

A Type II error is less harmful than a Type II error since Type II error only make us miss a real effect; A Type I error would make us have a misleading decisions based on a nonexistent relationship.

### Question 4

*1 point*

::: callout-note
1/1
:::

When conducting a difference-of-means test, which of the following samples would yield a sampling distribution with the lowest variability?

A.  A sample of 900 with a standard deviation of 15
B.  A sample of 25 with a standard deviation of 10
C.  A sample of 625 with a standard deviation of 20
D.  A sample of 100 with a standard deviation of 6

```{r}
15/sqrt(900)
10/sqrt(25)
20/sqrt(625)
6/sqrt(100)
```

Answer: A

### Question 5

*1 point*

::: callout-note
1/1
:::

Which of the following probabilities is not independent?

A.  The probability that the roulette wheel will end up on 23 two times in a row.
B.  The probability that three successive coin tosses will each turn up heads.
C.  The probability that I draw an ace and then a king in a row from a deck of cards (when drawing a two-card hand).
D.  The probability that I will get a 6 on three subsequent die rolls.
E.  None of the above – all are independent probabilities.

Answer: C

### Question 6

*1 point*

::: callout-note
1/1
:::

I conduct a two-tailed difference-of-means test and obtain a t-statistic of 2.10. Is my result statistically significant (with 95% confidence)?

A.  Yes, at the 0.05 level (p\<0.05)
B.  Yes, but only at the 0.10 level (p\<0.10)
C.  Not at either the 0.05 or the 0.10 level
D.  There is not enough information to answer this question.

Answer: D, we do not have sample size, so we do not know the DF.

### Question 7

*1 point*

::: callout-note
1/1
:::

I take a sample of 1800 adults and find that 360 of them watched last Monday’s NFL game. What probability represents the complement to the sample proportion of adults who watched the NFL game?

```{r}
1 - (360/1800)
```

Answer: 0.8

### Question 8

*1 point*

::: callout-note
0/1 Correct answer is A.
:::

Which of the following makes it more likely that a given sample statistic will be statistically different from zero (and thus allow you to reject the null hypothesis, all else equal)?

A.  More observations
B.  Greater variance in the sample
C.  Using a two-tailed instead of a one-tailed significance test
D.  A larger confidence interval around the test statistic
E.  Both (a) and (b) (but not (c) or (d))
F.  All of the above increase the chances of statistical significance
G.  None of the above

Answer: E, in both A and B, the changes will result in larger SE and smaller T.

### Question 9

*1 point*

::: callout-note
1/1
:::

The p-value for a two-tailed test of the null hypothesis $H_0: \mu=40$ is 0.06. Which of the following statements is accurate?

A.  A 95% confidence interval around the sample mean includes the value 40
B.  A 90% confidence interval around the sample mean includes the value 40
C.  A 99% confidence interval around the sample mean does not include the value 40
D.  None of the above statements are correct
E.  All of the above statements are correct

Answer: A \### Question 10

*4 points total*

#### Part A

*2 points*

::: callout-note
2/2
:::

If you roll a fair (six-sided) die twice, what is the probability that both rolls will be odd or greater than four?

```{r}
(1/2 + 1/3 - 1/6) ^ 2
```

Answer: 0.4444444 \#### Part B

*2 points*

::: callout-note
2/2
:::

The following is a distribution of U.S. college students classified by their age and full- vs. part-time status. Based on these data, what is the probability that a student is in the 25-29 age group, given that (i.e., conditional on knowledge that) the student is full time?

```{r}
tibble::tibble(
  age = c("15 - 19", "20 - 24", "25 - 29", "30+"),
  full_time = c(155, 255, 75, 35),
  part_time = c(20, 55, 80, 95)
) |> 
  knitr::kable()
```

```{r}
75 / (155 + 255 + 75 + 35)
```

Answer: 0.1442308, the probability is 14.4%. \### Question 11

*4 points total*

#### Part A

*2 points*

::: callout-note
2/2
:::

Using a SRS of 1211 people, I estimate that the proportion of people living in the South that favor teaching sexual education in public schools is 0.874 and the proportion of people outside of the south that favor teaching sexual education in public schools is 0.906. And, the standard error of the difference (in citizen views about teaching sexual education in public schools) between people living in the south and those not living in the south is 0.015. Give an interval estimate for the difference in the proportion of people favoring sex education in public schools between people who do, and do not, live in the south.

```{r}
(0.874 - 0.906) + 1.96 * 0.015 

(0.874 - 0.906) - 1.96 * 0.015 
```

Answer: A interval estimate for the difference in the proportion of people favoring sex education in public schools between people who do, and do not, live in the south is \[-0.0026, -0.0614\], so people live in south is 6.14% to 0.26 % lower in proportion favor teaching sexual education than people outside of the south.

#### Part B

*2 points*

::: callout-note
2/2
:::

Do the data (i.e., estimates above) show support for my hypothesis? How do you know?

```{r}
2*pnorm((0.874-0.906)/0.015, lower.tail = T)
```

With the p-value of 0.03289739, which is statistically significant in a two tail test, we can reject the null hypothesis. So the data shows support for the hypothesis.

### Question 12

*6 points total*

I am interested in estimating the average number of texts that University of Maryland undergraduate students send in a day. My hypothesis is that the average number of texts students send is greater than 100. I take a SRS of 1600 students and find that the mean number of texts they send is 105, and with a standard deviation of 120.

#### Part A

*2 points*

::: callout-note
1/2 You need to show how you calculate the critical t-statistic.
:::

What is a 95% confidence interval around the sample statistic?

```{r}
105 - 1.96 * (120/ sqrt(1600))
105 + 1.96 * (120/ sqrt(1600))
```

Answer: a 95% confidence interval around the sample statistic is \[99.12, 110.88\] \#### Part B

*2 points*

::: callout-note
1/2 You correctly identified the required test statistic (one-sample $t$-test), but you failed to show the calculation of the $t$-statistic itself. **You must show the formula and the substitution of values** ($t = \frac{105-100}{120/\sqrt{1600}}$) to earn full credit.
:::

When testing the null hypothesis, what is the test statistic associated with the sample statistic? Answer: We need an one-sample t-test statistic

#### Part C

*2 points*

::: callout-note
1/2 **You used `pnorm` (Z-distribution)** to calculate the $p$-value for a one-sample mean hypothesis test. Since the population standard deviation ($\sigma$) is unknown, you must use the **t-distribution** (`pt`) for the $p$-value. However, you drew the correct conclusion (reject $H_0$) based on the Z-test's $p$-value, so you are awarded follow-through credit.
:::

If using a one-tailed test of the null hypothesis and you are willing to accept a Type-I error rate of 0.05, do the data support my hypothesis? Why or why not?

```{r}
pnorm((105-100)/(120/sqrt(1600)), lower.tail = F)
```

With a p-value of 0.04779035, which is statistically significant, the one-tailed test will let us reject the null hypothesis. \## Applied questions

::: callout-note
All datasets referenced below are available through the `poliscidata` package.
:::

**Total points: 24**

### Question 1

```{r}

gss <- poliscidata::gss |> 
  mutate(voted08 = as.numeric(voted08),
         conarmy = as.numeric(conarmy))
```

*8 points total*

I hypothesize that, among only those that were eligible to vote, people with greater confidence in the U.S. military were more likely to turnout to vote in the 2008 presidential election. Use data from the General Social Survey (i.e., the `gss` dataset) to test my hypothesis (the unit of analysis is the individual survey respondent). Specifically, use the following variables: `conarmy` (1 = a “great deal” of confidence; 2 = “only some” confidence; 3 = “hardly any”); and `vote08` (1 = did not vote; 2 = voted). Answer the following questions.

#### Part A

*3 points*

::: callout-note
2/3 **Deduct 1 point** for an unclear and contradictory interpretation. Your final interpretation states that those with greater confidence were *more* likely to vote (74.2% \> 63.2%), which contradicts your preceding statement. **The table shows that** $74.2\%$ of those with 'a great deal' (conarmy=1) voted, and $63.2\%$ of those with 'hardly any' (conarmy=3) voted. Your interpretation should clearly state that *greater confidence* (lower `conarmy` value) is associated with a *higher* proportion of voting.
:::

Complete a cross-tab and interpret the results. Do the data support my hypothesis? Be sure to explain the nature of the relationship (or lack thereof, if relevant).

```{r}

datasummary_crosstab(voted08 ~ conarmy, data = gss |>
                       drop_na(voted08, conarmy),
                     statistic = 1~1 + N + Percent("col"))

```

Answer: I think the data support the hypothesis, there is a modest but clear relationship: respondents with lower confidence in the military were more likely to report voting in 2008.For those that were eligible to vote, people with greater confidence in the U.S. military were more likely(74.2% \> 63.2%) to turnout to vote in the 2008 presidential election.

#### Part B

*3 points*

::: callout-note
2/3 **Deduct 1 point** for an incorrect final $\chi^2$ value due to an initial data error (using the wrong total sample size, `nrow(gss)` instead of the $N$ for the filtered data). **The correct** $\chi^2$ is $\approx 6.01$ (with $\text{DF}=2$), but your calculated value is $6.467779$. You are awarded follow-through credit for showing the required steps for calculating expected counts and the sum of $\frac{(\text{Obs}-\text{Exp})^2}{\text{Exp}}$.
:::

Compute (by hand) the chi-square statistic to test the null hypothesis of no relationship between these two variables. Be sure to show your work.

```{r}
obs_values <- count(gss, voted08, conarmy, name = "obs_n")

army_total <- count(gss, conarmy, name = "army_total")

vote_total <- count(gss, voted08, name = "vote_total")

obs_exp_counts <- vote_total %>% 
  expand_grid(army_total) %>%  
  relocate(conarmy) %>%  
  mutate(exp_n = (vote_total * army_total) / nrow(gss))  %>% 
  left_join(obs_values, by = c("conarmy", "voted08"))

obs_exp_counts

chi_sq <- obs_exp_counts %>% 
  filter(!is.na(voted08)) %>% 
  filter(!is.na(conarmy)) %>% 
  mutate(diff = obs_n - exp_n,
         diff_2 = diff^2,
         diff_2_standard = diff_2 / exp_n)  %>%  
  summarise(chi_sq = sum(diff_2_standard)) %>%  
  pull()

chi_sq


```

Answer: the chi-square statistic is 6.467779.

#### Part C

*2 points*

::: callout-note
2/2
:::

Using the chi-square statistic that you computed in question 1(b), can you reject the null hypothesis of no relationship between these two variables with 95% confidence? Why, or why not?

```{r}
pchisq(chi_sq, df = 2, lower.tail = F)
```

Answer: The p-value of 0.03940395 shows that there is a statistical significance and we can reject the null hypothesis of no relationship between these two variables.

### Question 2

```{r}

nes <- poliscidata::nes |> 
  dplyr::mutate(envir_drill = factor(envir_drill, levels = c("1. Favor",
                                                             "3. Neither favor nor oppose",
                                                             "2. Oppose")),
                envir_drill = as.numeric(envir_drill),
                pid_x = as.numeric(pid_x),
                relig_pray = as.numeric(relig_pray),
                voted2008 = as.numeric(voted2008))
```

*8 points total*

I hypothesize that citizens who do not support increased U.S. offshore drilling are more conservative than those who do not. Use data from the `nes` dataset to test my hypothesis (the unit of analysis is the individual survey respondent). Specifically, use the following variables: `envir_drill` (1 = “favor”; 2 = “neither favor nor oppose”; 3 = “oppose”); and `pid_x` (higher values represent less liberalism, or more conservatism). Answer the following questions.

#### Part A

*2 points*

::: callout-note
2/2
:::

Using these data, what is the point estimate for the mean conservatism/liberalism score among those that oppose drilling? What is the point estimate for the mean conservatism/liberalism score among those that are in favor of drilling?

```{r}
nes_clean <- nes %>%
  filter(!is.na(envir_drill), !is.na(pid_x))

mean_results <- nes_clean %>%
  group_by(envir_drill) %>%
  summarize(mean_pid = mean(pid_x), n = n(), sd = sd(pid_x))
            
mean_results
```

Answer: The mean score among those that oppose drilling is 2.64 and The mean score among those that support drilling 4.18

#### Part B

*6 points*

::: callout-note
5/6 **Deduct 1 point** for not showing the full work for the standard error calculation. You only provided the final SE calculation and the $t$-statistic. **You must show the substituted values for the pooled standard error formula** (or the unpooled formula, but you should state the degrees of freedom for the $t$-test as required). You are awarded follow-through credit for calculating the $t$-statistic and drawing the correct conclusion (reject $H_0$).
:::

Evaluate the null hypothesis that there is no difference in the mean conservatism/liberalism score among those that approve vs. disapprove of drilling. Do the data support my hypothesis? Why or why not? Be sure to show all work necessary to answer the question by hand (i.e., you may only use R to the extent that is absolutely necessary to complete the question; otherwise, you must show how you would answer the question by hand).

```{r}
#Standard error of the difference
SEnes <- sqrt(1.689037 ^2 /1060 + 2.186845 ^2/3069	)

# T-statistic
tst <- (2.636792 - 4.179537) / SEnes	

#p-value
2 * pt(tst, df = 1060-1, lower.tail = T)
```

Answer: The p-value of 9.912213e-100 is so small that it has statistically significance, thus we can reject the null hypothesis that there is no difference in the mean conservatism/liberalism score among those that approve vs. disapprove of drilling. Thus data supports our hypothesis.

### Question 3

*8 points total*

I hypothesize that people who express that religion is important to them were more likely to turnout to vote in the 2008 presidential election. Use data from the `nes` dataset to test my hypothesis (the unit of analysis is the individual survey respondent). Specifically, use the following variables: `relig_pray`, which describes how often the respondent prays (1 = several times a day; 2 = once a day; 3 = a few times a week; 4 = once a week or less; 5 = never); and `voted2008` (1 = did not vote; 2= voted). Answer the following questions.

#### Part A

*2 points*

::: callout-note
2/2
:::

Using these data, what is the point estimate for the proportion of respondents that voted (i.e., turnout rate), among citizens expressing that religion is not important? What is the point estimate for the proportion of respondents that voted, among citizens expressing that religion matters a great deal?

```{r}
nes2 <- nes %>%
  mutate(relig_pray = as.character(relig_pray),
                voted2008 = as.character(voted2008)) %>% 
  filter(!is.na(relig_pray), !is.na(voted2008))

datasummary_crosstab(voted2008 ~ relig_pray, data = nes2,
                     statistic = 1~1 + N + Percent("col"))


```

Answer: If Religion importance is Not important, the Turnout estimate 72.4% If Religion importance is Matters a great deal, the Turnout estimate 82.4%

#### Part B

*6 points*

::: callout-note
5/6 **Deduct 1 point** for using the **unpooled** standard error (SE) formula, which is generally used for confidence intervals. **You must use the pooled SE formula** for a two-sample proportion hypothesis test to test the null hypothesis of no difference ($\hat{p}_{1} = \hat{p}_{2} = \hat{p}_{pooled}$). You correctly calculated the $Z$-statistic and drew the correct conclusion, so you are awarded follow-through credit.
:::

Evaluate the null hypothesis that there is no difference in the proportion of voters (i.e., turnout rate) among citizens expressing that religion is not important vs. those reporting that religion matters a great deal. Do the data support my hypothesis? Why or why not? Be sure to show all work necessary to answer the question by hand (i.e., you may only use R to the extent that is absolutely necessary to complete the question; otherwise, you must show how you would answer the question by hand).

```{r}
#Sample Proportion
p_hat <- (1445 + 687) / (1753 + 949)

#SE
SE2 <- sqrt(p_hat * (1 - p_hat) * (1/1753 + 1/949))

#Z-statistic
Zst <- (0.724 - 0.824)/ SE2

#p-value
2 * pnorm(Zst, lower.tail = T)
```

Answer: The p-value of 1.188036e-09 is so small that it has statistically significance, thus we can reject the null hypothesis that there is no difference in the proportion of voters among citizens expressing that religion is not important vs. those reporting that religion matters a great deal Thus data supports our hypothesis.
