---
title: "Problem Set 4"
subtitle: "Due date: 14 October"
format: 
  html:
    self-contained: true
toc: true
editor: visual
execute: 
  echo: true
---

Please upload your completed assignment to the ELMs course site (under the assignments menu) and to your class Github repository. You need to upload both your Quarto document (with R code, as needed) and its rendered output (either as a PDF or HTML file).

**Total points: 30**

::: callout-note
30/30
:::

## Question 1

*Points: 5*

::: callout-note
5/5
:::

Suppose I am playing poker with a regular deck of 52 cards. I have three cards in my hand, one of which is a king. What is the probability that the next two cards I draw will both be kings?

```{r}
3/49 * 2/48
```

The answer is 0.00255102.

## Question 2

*Points: 5*

::: callout-note
5/5
:::

A survey of 1,500 randomly selected individuals reveals that 46% of them have a college degree and that 31% of them have a gun in their house. If you were to assume that college graduation and gun ownership are independent of one another, what is the probability that an individual is both a college graduate and does not have a gun in their house?

```{r}
0.46* (1 - 0.31)
```

The answer is 0.3174

## Question 3

*Points: 5*

::: callout-note
5/5
:::

I roll a fair six-sided dice and I get a number greater than three. What is the probability that the next two dice rolls will also generate numbers greater than three?

```{r}

0.5*0.5
```

The answer is 0.25 \## Question 4

*Points: 10*

::: callout-note
10/10
:::

Assume that in the United States, 15% of people are smokers and, of smokers, 60% smoke at least one pack a day. Among those who smoke at least a pack a day, 60% will develop lung cancer at some point in their lives while, in the rest of the population (which includes smokers who smoke less than a pack a day and non-smokers), only 10% will. What is the probability that a person who develops lung cancer was a smoker who smoked at least a pack a day?

```{r}

(0.15*0.6*0.6)/ ((0.15*0.6*0.6) + ((1-0.15) + 0.15*0.4)* 0.1 )

```

The answer is 0.3724138 \## Question 5

*Points: 5*

::: callout-note
5/5
:::

Assume that in the world, 60% of countries are democracies and 80% of countries have ratified the Rome Statute for the International Criminal Court. Assume further that 90% of democracies have ratified the Rome Accords. What is the probability that a country is either a democracy or has ratified the Rome Accords?

```{r}
0.6+0.8-0.6*0.9


```

The answer is 0.86
